---
title: "Cluster analysis"
output: html_notebook
---

Chapter 1

* columns: features
* rows: observations

cluster analysis is EDA where observations are divided into meaningful groups that share common features

distance = 1 - similarity
dist(x, method = "euclidean")
method (arguments): "euclidean", "maximum", "manhattan", "canberra", "binary" or "minkowski"

```{r}
two_players <- data.frame(x = c(5,15), y=c(4,10))
library(ggplot2)
# Plot the positions of the players
ggplot(two_players, aes(x = x, y = y)) + 
  geom_point()
  # Assuming a 40x60 field
  # lims(x = c(-30,30), y = c(-20, 20))

# Split the players data frame into two observations
player1 <- two_players[1, ]
player2 <- two_players[2, ]

# Calculate and print their distance using the Euclidean Distance formula
player_distance <- sqrt( (player1$x - player2$x)^2 + (player1$y - player2$y)^2 )
player_distance

```


dist() function
```{r}
three_players <- rbind(two_players, c(0,20))
# Calculate the Distance Between two_players
dist_two_players <- dist(two_players)
dist_two_players

# Calculate the Distance Between three_players
dist_three_players <- dist(three_players)
dist_three_players

```

Importance of scale
standardization -> N(0,1)
https://stats.stackexchange.com/questions/10289/whats-the-difference-between-normalization-and-standardization
Normalization rescales to [0,1]
Standardization rescales to N(µ = 0, ??^2 = 1)

```{r}
library(datasets)
three_trees <- trees

# Calculate distance for three_trees 
dist_trees <- dist(three_trees)

# Scale three trees & calculate the distance  
scaled_three_trees <- scale(three_trees)
dist_scaled_trees <- dist(scaled_three_trees)

# Output the results of both Matrices
print('Without Scaling')
dist_trees

print('With Scaling')
dist_scaled_trees

```


categorical data

binary

* Jaccard Index, J(A,B) = intersection/union


dummification
```{r}
library(dummies)
# Dummify the Survey Data
dummy_survey <- dummy.data.frame(job_survey)

# Calculate the Distance
dist_survey <- dist(dummy_survey, method = "binary")

# Print the Original Data
job_survey

# Print the Distance Matrix
dist_survey

```

Chapter 2

* Complete linkage: maximum distance between two sets
* Single linkag: minimum
* Average linkage

```{r}
dist_players <- dist_three_players
# Extract the pair distances
distance_1_2 <- dist_players[1]
distance_1_3 <- dist_players[2]
distance_2_3 <- dist_players[3]

# Calculate the complete distance between group 1-2 and 3
complete <- max(c(distance_1_3, distance_2_3))
complete

# Calculate the single distance between group 1-2 and 3
single <- min(c(distance_1_3, distance_2_3))
single

# Calculate the average distance between group 1-2 and 3
average <- mean(c(distance_1_3, distance_2_3))
average


```

Hierarchical clustering
```{r}
pkgs <- c("dplyr", "ggplot2")
lapply(pkgs, library, character.only = TRUE)

players <- data.frame(x = c(-1, -2, 8, 7, -12, -15), y = c(1, -3, 6, -8, 8, 0))
dist_players <- dist(players, method = 'euclidean')
hc_players <- hclust(dist_players, method = 'complete')
cluster_assignments <- cutree(hc_players, k = 2)
players_clustered <- players %>%
  mutate(cluster = cluster_assignments)

ggplot(players_clustered, aes(x = x, y = y, color = factor(cluster))) +
  geom_point()

```

```{r}
lineup <- data.frame(x = c(-1, -2, 8,  7, -12, -15, -13, 15, 21,  12, -25, 26), 
                     y = c( 1, -3, 6, -8,   8,   0, -10, 16,  2, -15,   1,  0))

# Calculate the Distance
dist_players <- dist(lineup)

# Perform the hierarchical clustering using the complete linkage
hc_players <- hclust(dist_players, method = "complete")

# Calculate the assignment vector with a k of 2
clusters_k2 <- cutree(hc_players, k = 2)

# Create a new dataframe storing these results
lineup_k2_complete <- mutate(lineup, cluster = clusters_k2)

# Count the cluster assignments
count(lineup_k2_complete, cluster)

# Plot the positions of the players and color them using their cluster
ggplot(lineup_k2_complete, aes(x = x, y = y, color = factor(cluster))) +
  geom_point()
```

Dendrogram; aka, tree diagram

```{r}
# Prepare the Distance Matrix
dist_players <- dist(lineup)

# Generate hclust for complete, single & average linkage methods
hc_complete <- hclust(dist_players, method = "complete")
hc_single <- hclust(dist_players, method = "single")
hc_average <- hclust(dist_players, method = "average")

# Plot & Label the 3 Dendrograms Side-by-Side
# Hint: To see these Side-by-Side run the 4 lines together as one command
par(mfrow = c(1,3))
plot(hc_complete, main = 'Complete Linkage')
plot(hc_single, main = 'Single Linkage')
plot(hc_average, main = 'Average Linkage')
```

```{r}

library(dendextend)
dist_players <- dist(lineup, method = 'euclidean')
hc_players <- hclust(dist_players, method = "complete")

# Create a dendrogram object from the hclust variable
dend_players <- as.dendrogram(hc_players)

# Plot the dendrogram
plot(dend_players)

# Color branches by cluster formed from the cut at a height of 20 & plot
dend_20 <- color_branches(dend_players, h = 20)

# Plot the dendrogram with clusters colored below height 20
plot(dend_20)

# Color branches by cluster formed from the cut at a height of 40 & plot
dend_40 <-color_branches(dend_players, h = 40)

# Plot the dendrogram with clusters colored below height 40
plot(dend_40)


```

```{r}
dist_players <- dist(lineup, method = 'euclidean')
hc_players <- hclust(dist_players, method = "complete")

# Calculate the assignment vector with a h of 20
clusters_h20 <- cutree(hc_players, h = 20)

# Create a new dataframe storing these results
lineup_h20_complete <- mutate(lineup, cluster = clusters_h20)

# Calculate the assignment vector with a h of 40
clusters_h40 <- cutree(hc_players, h = 40)

# Create a new dataframe storing these results
lineup_h40_complete <- mutate(lineup, cluster = clusters_h40)

# Plot the positions of the players and color them using their cluster for height = 20
ggplot(lineup_h20_complete, aes(x = x, y = y, color = factor(cluster))) +
  geom_point()

# Plot the positions of the players and color them using their cluster for height = 40
ggplot(lineup_h40_complete, aes(x = x, y = y, color = factor(cluster))) +
  geom_point()

```

Wholesale dataset
```{r}
library(dendextend)
customers_spend <- readRDS("ws_customers.rds")
# Calculate Euclidean distance between customers
dist_customers <- dist(customers_spend)

# Generate a complete linkage analysis 
hc_customers <- hclust(dist_customers, method = "complete")

# Plot the dendrogram
plot(hc_customers)

# Create a cluster assignment vector at h = 15000
clust_customers <- cutree(hc_customers, h = 15000)

# Generate the segmented customers dataframe
segment_customers <- mutate(customers_spend, cluster = clust_customers)

dist_customers <- dist(customers_spend)
hc_customers <- hclust(dist_customers)
clust_customers <- cutree(hc_customers, h = 15000)
segment_customers <- mutate(customers_spend, cluster = clust_customers)

# Count the number of customers that fall into each cluster
count(segment_customers, cluster)

# Color the dendrogram based on the height cutoff
dend_customers <- as.dendrogram(hc_customers)
dend_colored <- color_branches(dend_customers, h = 15000)

# Plot the colored dendrogram
plot(dend_colored)

# Calculate the mean for each category
segment_customers %>% 
  group_by(cluster) %>% 
  summarise_all(funs(mean(.)))
```

Chapter 3

K-means cluster

soccer, we know k = 2. Initialize 2 centroids (random)

```{r}
lineup <- data.frame(x = c(-1, -2, 8,  7, -12, -15, -13, 15, 21,  12, -25, 26), 
                     y = c( 1, -3, 6, -8,   8,   0, -10, 16,  2, -15,   1,  0))

# Build a kmeans model
model_km2 <- kmeans(lineup, centers = 2)

# Extract the cluster assignment vector from the kmeans model
clust_km2 <- model_km2$cluster

# Create a new data frame appending the cluster assignment
lineup_km2 <- mutate(lineup, cluster = clust_km2)

# Plot the positions of the players and color them using their cluster
ggplot(lineup_km2, aes(x = x, y = y, color = factor(cluster))) +
  geom_point()

```

```{r}

# Build a kmeans model
model_km3 <- kmeans(lineup, centers =3)

# Extract the cluster assignment vector from the kmeans model
clust_km3 <- model_km3$cluster

# Create a new data frame appending the cluster assignment
lineup_km3 <- mutate(lineup, cluster = clust_km3)

# Plot the positions of the players and color them using their cluster
ggplot(lineup_km3, aes(x = x, y = y, color = factor(cluster))) +
  geom_point()

```

```{r}
library(purrr)
lineup <- data.frame(x = c(-1, -2, 8,  7, -12, -15, -13, 15, 21,  12, -25, 26), 
                     y = c( 1, -3, 6, -8,   8,   0, -10, 16,  2, -15,   1,  0))

tot_withinss <- map_dbl(1:10,  function(k){
  model <- kmeans(x = lineup, centers = k)
  model$tot.withinss
})

# Generate a data frame containing both k and tot_withinss
elbow_df <- data.frame(
  k = 1:10 ,
  tot_withinss = tot_withinss
)

# Plot the elbow plot
ggplot(elbow_df, aes(x = k, y = tot_withinss)) +
  geom_line() +
  scale_x_continuous(breaks = 1:10)

```

