---
title: "EDA-bray"
output: html_notebook
---

## 1. Exploring categorical variables

levels() returns factor levels 

table() returns contingency table

```{r}
list.files()
comics <- read.csv("comics.csv")

# Check levels of align
levels(comics$align)

# Check the levels of gender
levels(comics$gender)

# Create a 2-way contingency table
tab <- table(comics$align, comics$gender)
```

dropping levels
```{r}
library(dplyr)

# Remove align level
comics <- comics %>%
  filter(align != "Reformed Criminals") %>%
  droplevels()
  

```

side-by-side barcharts

```{r}
# Load ggplot2
library(ggplot2)

# Create side-by-side barchart of gender by alignment
ggplot(comics, aes(x = align, fill = gender)) + 
  geom_bar(position = "dodge")

# Create side-by-side barchart of alignment by gender
ggplot(comics, aes(x = gender, fill = align)) + 
  geom_bar(position = "dodge") +
  theme(axis.text.x = element_text(angle = 90))
```

counts vs proportions

prop.table() proportions

sum(prop.table()) equals one

prop.table( , 1 or 2) conditions or rows or columns

```{r}
tab <- table(comics$align, comics$gender)
options(scipen = 999, digits = 3) # Print fewer digits
prop.table(tab)     # Joint proportions
prop.table(tab, 2)  # Conditional on columns
```

```{r}
# Plot of gender by align
ggplot(comics, aes(x = align, fill = gender)) +
  geom_bar() # count by defaul
  
# Plot proportion of gender, conditional on align
ggplot(comics, aes(x = align, fill = gender)) + 
  geom_bar(position = "fill") +
  ylab("proportion")
```

Distribution of one variable

filter or facet to condition 
tilde (~) can be read as "broken down by"

generally bar charts > pie charts

```{r}
# Change the order of the levels in align
comics$align <- factor(comics$align, 
                       levels = c("Bad", "Neutral", "Good"))

# Create plot of align
ggplot(comics, aes(x = align)) + 
  geom_bar()
```




```{r}
# Plot of alignment broken down by gender
ggplot(comics, aes(x = align)) + 
  geom_bar() +
  facet_wrap(~ gender)
```

improve piechart
```{r}
# Put levels of flavor in decending order
lev <- c("apple", "key lime", "boston creme", "blueberry", "cherry", "pumpkin", "strawberry")
pies$flavor <- factor(pies$flavor, levels = lev)

# Create barchart of flavor
ggplot(pies, aes(x = flavor)) + 
  geom_bar(fill = "chartreuse") + 
  theme(axis.text.x = element_text(angle = 90))
```

## 2. Exploring numerical data

dotplot zero information loss
histogram aggregates with binning (some information loss)
density plot for larger sets


```{r}
library(ggplot2)
cars <- read.csv("cars04.csv")
str(cars)
```

```{r}
# Learn data structure
str(cars)

# Create faceted histogram
ggplot(cars, aes(x = city_mpg)) +
  geom_histogram() +
  facet_wrap(~ suv)
```

```{r}
# Filter cars with 4, 6, 8 cylinders
common_cyl <- filter(cars, ncyl %in% c(4,6,8))

# Create side-by-side box plots of city mpg by ncyl
ggplot(common_cyl, aes(x = as.factor(ncyl), y = city_mpg)) +
  geom_boxplot()

# Create overlaid density plots for same data
ggplot(common_cyl, aes(x = city_mpg, fill = as.factor(ncyl))) +
  geom_density(alpha = .3)
```

Distribution of one variable

marginal and conditional histograms
```{r}
# Create hist of horsepwr
cars %>%
  ggplot(aes(x = horsepwr)) +
  geom_histogram() +
  ggtitle("Distribution of horsepower for all cars")

# Create hist of horsepwr for affordable cars
cars %>%
  filter(msrp < 25000) %>%
  ggplot(aes(x = horsepwr)) +
  geom_histogram() +
  xlim(c(90, 550)) +
  ggtitle("Distribution of horsepower for affordable cars")
```

Box plots

box plot based on 1st, 2nd (median), and 3rd quartile

whiskers incorportes "almost all of the data" except identified outliers

weakness of box plot is cannot identify multiple modes


```{r}
# Construct box plot of msrp
cars %>%
  ggplot(aes(x = 1, y = msrp)) +
  geom_boxplot()

# Exclude outliers from data
cars_no_out <- cars %>%
  filter(msrp < 100000)

# Construct box plot of msrp using the reduced dataset
cars_no_out %>%
  ggplot(aes(x = 1, y = msrp)) +
  geom_boxplot()
  
```


Plot selection

"Because the city_mpg variable has a much wider range with its outliers, it's best to display its distribution as a box plot."
```{r}
# Create plot of city_mpg
cars %>%
  ggplot(aes(x = 1, y = city_mpg)) +
  geom_boxplot()

# Create plot of width
cars %>% 
  ggplot(aes(x = width)) +
  geom_density()
```


Visualizing in higher dimensions
```{r}
# Facet hists using hwy mileage and ncyl
common_cyl %>%
  ggplot(aes(x = hwy_mpg)) +
  geom_histogram() +
  facet_grid(ncyl ~ suv) +
  ggtitle("Faceted on ncy and suv")
```


## 3. Numerical Summaries
```{r}
library(gapminder); library(dplyr)
data(gapminder)
str(gapminder)
glimpse(gapminder)
```


Calculate center measures
```{r}
library(ggplot2)
# Create dataset of 2007 data
gap2007 <- filter(gapminder, year == 2007)

# Compute groupwise mean and median lifeExp
gap2007 %>%
  group_by(continent) %>%
  summarize(mean(lifeExp),
            median(lifeExp))

# Generate box plots of lifeExp for each continent
gap2007 %>%
  ggplot(aes(x = continent, y = lifeExp)) +
  geom_boxplot()

```

Measures of variability

Calculate spread measures
```{r}
# Compute groupwise measures of spread
gap2007 %>%
  group_by(continent) %>%
  summarize(sd(lifeExp),
            IQR(lifeExp),
            n())

# Generate overlaid density plots
gap2007 %>%
  ggplot(aes(x = lifeExp, fill = continent)) +
  geom_density(alpha = 0.3)

```
```{r}
# Compute stats for lifeExp in Americas
gap2007 %>%
  filter(continent == "Americas") %>%
  summarize(mean(lifeExp),
            sd(lifeExp))

# Compute stats for population
gap2007 %>%
  summarize(median(pop),
            IQR(pop))
```

Shape and transformations

Center and spread; third is modality and skew

to treat the high skew: log transform

```{r}
# Create density plot of old variable
gap2007 %>%
  ggplot(aes(x = pop)) +
  geom_density()

# Transform the skewed pop variable
gap2007 <- gap2007 %>%
  mutate(log_pop = log(pop))

# Create density plot of new variable
gap2007 %>%
  ggplot(aes(x = log_pop)) +
  geom_density()
```

Outliers

```{r}
# Filter for Asia, add column indicating outliers
gap_asia <- gap2007 %>%
  filter(continent == "Asia") %>%
  mutate(is_outlier = lifeExp < 50)

# Remove outliers, create box plot of lifeExp
gap_asia %>%
  filter(!is_outlier) %>%
  ggplot(aes(x = 1, y = lifeExp)) +
  geom_boxplot()
```

## 4. Case study
```{r}

# Load packages
library(ggplot2)
library(dplyr)
library(openintro)


# Compute summary statistics
email %>%
  group_by(spam) %>%
  summarize(median(num_char),
    IQR(num_char))
  
# Create plot
email %>%
  mutate(log_num_char = log(num_char)) %>%
  ggplot(aes(x = spam, y = log_num_char)) +
  geom_boxplot()
```

Spam and !!!
```{r}

# Compute center and spread for exclaim_mess by spam
email %>%
  group_by(spam) %>%
  summarize(median(exclaim_mess),
            IQR(exclaim_mess))

# Create plot for spam and exclaim_mess
email %>%
  mutate(log_exclaim_mess = log(exclaim_mess + .01)) %>%
  ggplot(aes(x = log_exclaim_mess)) +
  geom_histogram() +
  facet_wrap(~ spam)

# Alternative plot: side-by-side box plots
email %>%
  mutate(log_exclaim_mess = log(exclaim_mess + .01)) %>%
  ggplot(aes(x = 1, y = log_exclaim_mess)) +
  geom_boxplot() +
  facet_wrap(~ spam)

# Alternative plot: Overlaid density plots
email %>%
  mutate(log_exclaim_mess = log(exclaim_mess + .01)) %>%
  ggplot(aes(x = log_exclaim_mess, fill = spam)) +
  geom_density(alpha = 0.3)

```


Zero inflation strategies

```{r}
# Create plot of proportion of spam by image

email %>%
  mutate(has_image = image > 0) %>%
  ggplot(aes(x = has_image, fill = spam)) +
  geom_bar(position = "fill")

```

```{r}
# Question 1
email %>%
  filter(dollar > 0 ) %>%
  group_by(spam) %>%
  summarize(median(dollar))

# Question 2
email %>%
  filter(dollar > 10 ) %>%
  ggplot(aes(x = spam)) +
  geom_bar()
```

What's in a number?

```{r}
# Reorder levels
email$number <- factor(email$number, levels = c("none", "small", "big"))

# Construct plot of number
ggplot(email, aes(x = number)) + 
  geom_bar() +
  facet_wrap(~spam)
  
```

