---
title: "notes: dc/writing functions"
output: html_notebook
---

## 1. Refresher; function template
## 2. When and how should you write a function?
## 3. Why functional programming?
## 4. Dealing with failure

```{r}
setwd("D:/GitHub/dc/write-func")
```


## 1. Refresher; function template

my_fun <- function(arg1, arg2) {
  body
}

The last expression evaluated in a function is the return value

Functions fundamentals

1. Formal arguments
2. Body
3. Environment

```{r}
add <- function(x, y=1) {
  x+y
}
formals(add)
body(add)
environment(add)
```

My Black-Scholes
```{r}

BlackScholes_call <- function(S, K, r, sigma, T) {
  d1 <- (log(S/K) + (r + sigma^2/2)*T)/(sigma*sqrt(T))
  d2 <- d1 - sigma*sqrt(T)
  Nd1 <- pnorm(d1)
  Nd2 <- pnorm(d2)
  c <- S*Nd1 - K*exp(-r*T)*Nd2
  c
}
```


Environments
Scoping describes WHERE, not when, to look for a value

* When you call a function, a new environment is made for the function
* The new environment is populated with the argument values
* Objects are looked for first in this environment
* If they are not found, they are looked for in the environment that the function was created in

```{r}
# 1
y <- 10
f <- function(x) {
  x + y 
}
f(10)
# returns 20; because y is not passed as an argument to the function, R looks outside of the function environment

# 2
y <- 10
f <- function(x) {
  y <- 5
  x + y
}
f(10)
#returns 15; the value of y is defined inside of the function

# 3
f <- function(x) {
  y <- 5
  x + y
}
f(5)
# what is value of y the variable? 
# Answer: object 'y' not found; i.e., does not exist in the global environment

```


### Data structures

Two types of vectors in R: atomic, lists.

* Atomic vectors: logical, integer, double, character, complex, raw
* Lists; aka recursive vectors can contain heterogeneous objects

Every vector has two key properties

* typeof()
* length()

Missing values

* NULL indicates absence of a vector
* NA aka missing value; NA is a missing element inside a vector
* is.na() returns a logical
* missing values are contagious

### Lists

Contain heterogeneous objects
Created with list()

* Single [ extracts sublist
* [[ and $ extract elements


### for loops

seq_along() generates a sequence along the index of the object passed to it


```{r}

# Arguments
# Rewrite the call to follow best practices
mean(0.1,x=c(1:9, NA),TRUE)
mean(x = c(1:9, NA), trim = 0.1, na.rm = TRUE)
mean(c(1:9, NA), trim = 0.1, na.rm = TRUE)

f <- function(x) {
  if (TRUE) {
    return(x + 1)
  }
  x
}

y <- 10
f <- function(x) {
  print(environment(y))
  x + y
}
environment(f)
environment(y)
f(10)
# Correct! Because y is not passed in as an argument to the function, R looks outside of the function environment.

y <- 10
f <- function(x) {
  y <- 5
  x + y
}


df <- data.frame(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

df <- data.frame()
1:ncol(df)

for (i in 1:ncol(df)) {
  print(median(df[[i]]))
}


# Replace the 1:ncol(df) sequence
for (i  in seq_along(df)) {
  print(median(df[[i]]))
}

# Change the value of df
df <- data.frame()

# Repeat for loop to verify there is no error
for (i in seq_along(df)) {
  print(median(df[[i]]))
}


```

## 2. When and how should you write a function?


```{r}
# Define example vector x
x <- 1:10 

# Use the function template to create the rescale01 function
rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2]- rng[1])
}

rescale01(x)
```

### How should you write a function?

Start with a simple problem,
get a working code snippet
rewrite using temp variables
rewrite for clarity
finally, turn snippet into function

```{r}
# Define example vectors x and y
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3,  4)

# Turn this snippet into a function: both_na()
both_na <- function(x, y) {
  sum(is.na(x) & is.na(y))
}
```

### How can you write a good function?

Should be both correct and understandable

Naming principles

* use verbs for function names; e.g., add_row()

Exercise
This is a case where an early return() makes sense. If x is empty, the function should immediately return c(-Inf, Inf).

```{r}

mean_ci <- function(x, level = 0.95) {
  if (length(x) == 0) {
    warning("`x` was empty", call. = FALSE)
    return(c(-Inf, Inf))
  }
  se <- sd(x) / sqrt(length(x))
  alpha <- 1 - level
  mean(x) + se * qnorm(c(alpha / 2, 1 - alpha / 2))
}

### Next function is improved as follows

replace_missings <- function(x, replacement) {
  is_miss <- is.na(x)
  x[is_miss] <- replacement
  
  # Rewrite to use message()
  message(paste(sum(is_miss),"missings replaced by the value", replacement))
  x
}

# Check your new function by running on df$z
df$z <- replace_missings(df$z, 0)

```


## 3. Why functional programming?

purr package

```{r}

df <- data.frame(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

# Initialize output vector
output <- vector("double", ncol(df))  

# Fill in the body of the for loop
for (i in seq_along(df)) {            
  output[[i]] <- median(df[[i]], na.rm=TRUE)
}

# View the result
output

# But this can be a function

col_median <- function(df) {
  output <- vector("double", ncol(df))
  for (i in seq_along(df)) {
    output[[i]] <- median(df[[i]])  # df[[i]] returns the ith column as numeric vector; df[i] returns the ith column as df
  }
  output
}


# Add a second argument called power
f <- function(x, power) {
  # Edit the body to return absolute deviations raised to power
  abs(x - mean(x))^power
}

```

### Functions can be arguments, too

```{r}

# this is essentiallly our own sapply()
col_summary <- function(df, fun) {
  output <- vector("numeric", length(df))
  for (i in seq_along(df)) {
    output[[i]] <- fun(df[[i]])
  }
  output
}

col_summary(df, median)


```

### Introducing purrr

map(.x, .f, ...)

The map functions

All the map functions in purrr take a vector, .x, as the first argument, then return .f applied to each element of .x. The type of object that is returned is determined by function suffix (the part after _):

* map() returns a list or data frame
* map_lgl() returns a logical vector
* map_int() returns a integer vector
* map_dbl() returns a double vector
* map_chr() returns a character vector

```{r}
# library(purrr)

data("mtcars")
cyl <- split(mtcars, mtcars$cyl)

# Examine the structure of cyl
str(cyl)

# Extract the first element into four_cyls
four_cyls <- cyl[[1]]

# Fit a linear regression of mpg on wt using four_cyls
lm(mpg ~ wt, four_cyls)

# Rewrite to call an anonymous function
map(cyl, function(df) lm(mpg ~ wt, data = df))

# Rewrite to use the formula shortcut instead
# map(cyl, function(df) lm(mpg ~ wt, data = df))
map(cyl, ~ lm(mpg ~ wt, data = .))

# Save the result from the previous exercise to the variable models
models <- map(cyl, ~ lm(mpg ~ wt, data = .))

# Use map and coef to get the coefficients for each model: coefs
coefs <- map(models, coef)

# Use string shortcut to extract the wt coefficient 
map(coefs, "wt")


# with pipes

# Define models (don't change)
models <- mtcars %>% 
  split(mtcars$cyl) %>%
  map(~ lm(mpg ~ wt, data = .))

# Rewrite to be a single command using pipes 
models %>% map(summary) %>% map_dbl("r.squared")

```

## 4. Dealing with failure

purrr package: 

* safely() "capture side effects" and plays nicely with map() functions
* possibly()
* quietly()

```{r}
library(purrr)
map(list(5,10,20), rnorm)

# map2() to iterate over two arguments
map2(list(5,10,20), list(1, 5, 10), rnorm)

# pmap() to iterate over many arguments
print("this is pmap ....")
pmap(list(n = list(5, 10, 20),
          mean = list(1, 5, 10), 
          sd = list(0.1, 0.5, 0.1)), rnorm)

# invoke_map() to iterate over functions
invoke_map(list(rnorm, runif, rexp), n = 7)

```

Exercises
```{r}
# Initialize n and mu
n <- list(5, 10, 20)
mu <- list(1, 5, 10)

# Create a sd list with the values: 0.1, 1 and 0.1
sd <- list(.1, 1, .1)

# Edit this call to pmap() to iterate over the sd list as well
pmap(list(n, mu, sd), rnorm)
```

```{r}
# Define list of functions
f <- list("rnorm", "runif", "rexp")

# Parameter list for rnorm()
rnorm_params <- list(mean = 10)

# Add a min element with value 0 and max element with value 5
runif_params <- list(min = 0, max = 5)

# Add a rate element with value 5
rexp_params <- list(rate = 5)

# Define params for each function
params <- list(
  rnorm_params,
  runif_params,
  rexp_params
)

# Call invoke_map() on f supplying params as the second argument
invoke_map(f, n = 5, params)
```

walk() operates just like map() except it's designed for functions that don't return anything. You use walk() for functions with side effects like printing, plotting or saving.

```{r}
# Define list of functions
f <- list(Normal = "rnorm", Uniform = "runif", Exp = "rexp")

# Define params
params <- list(
  Normal = list(mean = 10),
  Uniform = list(min = 0, max = 5),
  Exp = list(rate = 5)
)

# Assign the simulated samples to sims
sims <- invoke_map(f, params, n = 50)

# Use walk() to make a histogram of each element in sims
walk(sims, hist)
```

```{r}
# Define list of functions
f <- list(Normal = "rnorm", Uniform = "runif", Exp = "rexp")

# Define params
params <- list(
  Normal = list(mean = 10),
  Uniform = list(min = 0, max = 5),
  Exp = list(rate = 5)
)

# Assign the simulated samples to sims
sims <- invoke_map(f, params, n = 50)


breaks_list <- list(
  Normal = seq(6, 16, 0.5),
  Uniform = seq(0, 5, 0.25),
  Exp = seq(0, 1.5, 0.1)
)

# Use walk2() to make histograms with the right breaks
walk2(sims, breaks_list, hist)

```


## 5. Robust functions

```{r}
options() # global options
getOption("digits")

```

